### 2015-07-14 paper/main.tex
change bibliography
Fix typos, minor edits

While these new models of international trade are well grounded in empirical evidence coming from micro data, there are surprisingly few tests of the model predictions for aggregate variables which are decisive for the predicted welfare gains from trade. Recently, \citet{Arkolakisetal2012a,Arkolakisetal2012b} call into question the importance of firm-level heterogeneity by showing that in a lass class of trade models, the additional welfare gains are fairly small and actually even smaller if consumers don't have CES utility. The response of \citet{MelitzRedding2013} shows that there is still considerable disagreement over how to theoretically evaluate the additional welfare gains from firm selection, a fact that motivates us to test the model directly in aggregate data on prices, markups and productivity. To do so, we estimate the effects of trade liberalization on the competitive environment in manufacturing markets of the member countries of the North American Free Trade Agreement (NAFTA). We employ an estimation procedure based on the \citet{MelitzOttaviano2008} model introduced by \citet{Chen2009}, which to our knowledge is the only empirical application of a model with firm-level heterogeneity on aggregate data. \citet{Chen2009} derive estimable regression equations from the model's equilibrium conditions that allow us to test the effects of trade openness on relative price levels, markups and labor productivities of two trading partners. It is further possible to differentiate between the effects of trade in the short run, which, in the model, refers to an economy without relocation decisions for firms, and in the long run, when firms are free to choose their home market for production. However, as the underlying model is static, no direct results on the time path of the impact of trade liberalization can be obtained. We try to address this issue by dividing our sample in ways that make it more amenable to a model-based estimation. Contrary to \citet{Chen2009}, we directly observe tariff rates between the three countries in our sample and hence use those as a direct measure of trade openness. Additionally, we test for the effects of third-country trade openness on the relative performance of two countries that are linked through trade, predictions for which can be derived from the multi-country version of the Melitz and Ottaviano model. Our dataset comprises of nine manufacturing sectors in Canada, Mexico and the US, covering the time period from the introduction of NAFTA in 1994 up to 2006, which gives us reason to believe that we are able to capture the long run effects of policy changes even in industries with low firm churning rates.\footnote{Given that the US already engaged in a free-trade agreement with Canada in 1988 (CUSFTA), we also estimate the short-run effects of this agreement prior to NAFTA and the results hold}. \\
Our findings support the main model predictions, with tariff barriers stifling domestic competition, leading to higher producer prices, markups and lower productivity. In the immediate years after the free-trade agreement when tariff barriers are reduced, relative prices and markups decrease as relative productivity increases, thus giving rise to competitive effects. The results in the long-run, however, are not as clear cut, with some effects reversing as predicted by the model while some effects persist. This is also confirmed by directly looking at the reaction of industries with different entry barriers to changes in trade openness.  

@@

The paper is organized as follows: Section \ref{sec:lit} gives a survey of previous literature assessing the effects of trade liberalizations in general and of NAFTA specifically. Section \ref{sec:mo} briefly summarizes the Melitz and Ottaviano (2008) model, derives the most important equilibrium conditions and explains the estimation strategy used in \citet{Chen2009}. Section \ref{sec:app} then presents our application of the model by giving an overview of the data used and our estimation procedure. The results of our regressions and possible shortcomings as well as extensions of our approach are discussed in Section \ref{sec:disc}; Section \ref{sec:conc} concludes.

@@

c_D = \left[ \frac{\phi c_M^k}{L} \frac{1-\tau^{-k}}{1-(\tau\tau^{*})^{-k}} \right]^{\frac{1}{k+2}}
c_D^* = \left[ \frac{\phi c_M^k}{L} \frac{1-\tau^{-k}}{1-(\tau\tau^{*})^{-k}} \right]^{\frac{1}{k+2}},

@@

\left( \frac{\bar{p}}{\bar{p}^*} \right)^k = \left( \frac{c_D}{c_D^*} \right)^k = \left(\frac{c_M}{c_M^*} \right)^k \frac{\bar{N}*}{\bar{N}} \frac{N}{N^*} \frac{1 + \frac{\bar{N}}{\bar{N}^*} \frac{1}{(\tau^*)^k} \left( \frac{c_M^*}{c_M} \right)^k}{1 + \frac{\bar{N}^*}{\bar{N}} \frac{1}{\tau^k} \left( \frac{c_M}{c_M^*} \right)^k} \label{eq:mo-relprice-sr}

@@

\left( \frac{\bar{p}}{\bar{p}^*} \right)^(k+2) = \left( \frac{c_D}{c_D^*} \right)^(k+2) = \left( \frac{c_M}{c_M^*} \right)^k \frac{L^*}{L} \frac{1-\frac{1}{\tau^k}}{1 - \frac{1}{(\tau^*)^k}} \label{eq:mo-relprice-lr}

@@

These two equations capture one of the central predictions of the Melitz and Ottaviano model: asymmetrical trade liberalizations will have opposing effects on competitiveness in the short and the long run. By equation \ref{eq:mo-relprice-sr}, lowering trade barriers induces a fall in the cost cutoff, and hence decreases in prices and markups and increases in productivity. In the long run, however, the effects are reversed, as an increase in trade costs induces firms to choose the relatively more protected market for production, thereby increasing competition in markets that are shielded from foreign firms. \\

@@

Similarly, an expression for the average markup can be derived. The determination of the average markup is equivalent to the one for average prices so expressions for the short-- and long--run impacts of openness on markups can readily be derived. Somewhat more problematic is the index for productivity, as the model requires knowledge of a firm's unit costs $c$, which are not observable. Chen et al. work around this issue by assuming away differences in capital costs, so that average industry productivity can be approximated by the ratio of nominal wages to labor productivity: $\bar{c} = \frac{w}{z}$. If it is additionally assumed that unit labor costs only depend on nominal wages, the ratio of domestic to foreign labor productivity can be written as:
\begin{equation}\label{eq:chen-lab-prod}
\frac{z}{z^*} = \frac{w}{w^*} \frac{\bar{c}^*}{\bar{c}}
\end{equation}
If the least competitive firm in an industry with a productivity draw at the upper bound of the distribution $c_M$ has labor productivity $z_M$ and labor is perfectly mobile between firms, equation ??? implies $\frac{z}{z^*} = \frac{w}{w^*} \frac{c_M^*}{c_M}$. This relationship can then be used in an analogous fashion as before to construct an expression relating openness to productivity. In the short run, equation ??? can be amended to yield:
\begin{equation}\label{eq:chen-prod-sr}
\left( \frac{z}{z^*} \right)^k = \left( \frac{z_M}{z_M^*} \right)^k \frac{(\bar{N}/N)}{(\bar{N}^*/N^*)} \frac{1+ \frac{\bar{N}*}{\bar{N}} \frac{\theta}{1-\theta}}{1+ \frac{\bar{N}}{\bar{N}^*} \frac{\theta^*}{1-\theta^*}} 
\end{equation}
Higher values of $\theta$ thus lead to higher productivity (conditional on $\bar{N}/N$), as they force lower productivity firms to shut down production. For the long run, equation ??? combined with the expression for labor productivity gives:
\begin{equation}\label{eq:chen-prod-lr}
\left( \frac{z}{z*} \right)^{k+2} = \left( \frac{w}{w^*} \right)^2 \frac{L}{L^*} \left( \frac{z_M}{z_M^*} \right)^k \frac{1-\frac{\theta}{1-\theta}}{1-\frac{\theta^*}{1-\theta^*}}
\end{equation} 
Larger markets exhibit higher labor productivity, while the effects of $\theta$ and $\theta^*$ are the opposite of those in the short--run.

@@

\subsection{Further Changes (work in progress!)}
%Following the arguments in \citet{Trefler2004}, we want to exploit the nature of NAFTA being close to a natural experiment and hence try to identify the effect of the policy measures (i.e. the changes in tariff rates) separately from the effects of trade openness in general. There is evidence that trade openness -- measured by the import penetration of a certain country or industry, as in \citet{Chen2009} -- is affected by a number of external forces, including oil prices \citep{Kilian2009}, credit conditions \citep{Chor2012}, shared culture and language between countries and many more. Therefore, we deconstruct the iceberg costs of trade into two parts: $\tau^{lh}=\frac{Tr^{lh}}{\theta^{lh}}$, where $Tr^{lh}>1$ is the tariff rate for trade between countries $l$ and $h$ and $\theta^{lh}>1$ captures the additional costs of trade imposed by the aforementioned factors. Then, the import penetration can be viewed as a measure of $\theta^{lh}$, as in Chen et al. (2009), while a carefully constructed tariff measure can directly account for the effect of NAFTA-mandated changes in the trade environment. Obviously, the import penetration in a sector depends on the tariff measure as well, so an instrumental variable approach has to be taken in order to identify this effect. We defer the discussion of the construction of the tariff measure and the choice of instruments to the following chapter. Prior work of \citet{Bernard2006} suggests that tariff rates throughout the 1980s, at an average level between four and five percent, accounted for about the same fraction of trade costs as costs directly attached to shipping the good, i.e. freight and insurance, so we expect them to have a sizable impact on trade flows between countries and hence the competitive environment. 
As we have seen in the exposition of the Melitz and Ottaviano model above, there is one crucial caveat in taking the model to the data: due to the static nature of the model, the comparative static results just compare one steady state with another, while being silent about the transitional dynamics. The estimation strategy of \citet{Chen2009} tries to account for this by estimating an error correction model to identify the long-run separately from the short-run, but their results -- just as ours -- are mixed for the long run and it cannot be ruled out that this is due to the estimation procedure. Therefore, we try to address this issue in a more direct way: as short- and long-run in the model differ only in the entry, we separate industries into those with a fixed number of firms and those with low entry barriers. This distinction then gives us industries that represent the short- and long-run and we can directly investigate whether the coefficients on the relevant variables differ significantly\footnote{We were inspired to do so by \citet{Head1999} who use the classification to test competing theories of trade that rely on different market structures.}. This approach, however, leads to two issues that need to be addressed before implementation. First, it is not \textit{a priori} obvious how to measure the entry conditions in an industry; while the theoretical model uses the number of firms, this could in practice either refer to firms or to establishments (i.e. different production sites run by the same parent company), or even to employees, as firms in the model use unit labor input. Second, there is no reason to believe that different measures of entry and exit dynamics are exogenous with respect to trade openness -- indeed in the model trade openness is a key factor in the entry decision of firms, but in the real world there might be various other factors that might lead to industries being asymmetrically affected by a change in trade costs, hence biasing our results. To tackle both these issues, we aim to construct a robust measure of industry dynamics by aggregating multiple studies that examine firm and employment turnover in Canada, Mexico and the United States as well as Europe over different time periods. With this, we hope to identify those industries that are either very dynamic or very static over a broad set of different measures, regions and time periods. Table \ref{tb:market_lit} gives an overview of the studies used and a glance at their respective results, showing considerable variation in the dynamics of entry and job creation in different manufacturing sectors.

\begin{center}
\begin{table}\label{tb:market_lit}
\begin{tabular}{|P{2.8cm}|P{4cm}|P{4cm}|P{4cm}|}
\hline
\textbf{Study}  &  \textbf{Subject}  &  \textbf{Highest Turnover}  & \textbf{Lowest Turnover} \\
\hline 
\hline
\citet{Dunne1988} & Entry Rates (4-yearly) \newline U.S. \newline 63-82    & Instruments (.6) \newline Lumber (.5) \newline Printing (.49)  & Leather (.29) \newline  Food Processing (.24) \newline Tobacco (.21) \\  
\hline
Samianego (2008)   & Entry Rates (yearly)  \newline Europe \newline 97-04  & Paper, printing, software (15.6) \newline Textiles (11.9) \newline Petroleum and Coal (11.9) & Chemicals (9.5) \newline Plastics (9.4) \newline Food Products (9.1) \\
\hline 
\citet{Brown2004}  & Employment renewal    \newline Canada \newline 73-96  & Plastic (79.5) \newline Furniture (79.4) \newline Fabricated Metals (77.2) & Primary Metals (33.6) \newline Paper (32.4) \newline Tobacco (4.2) \\
\hline
\citet{Foster2006} & Job creation (yearly) \newline U.S. \newline 72-98     & Lumber (11.8) \newline Apparel (11.2) \newline Miscellaneous (11.0) & Paper (5.9) \newline Petroleum (5.9) \newline Tobacco (5.1) \\
\hline
\citet{Baldwin1994} & Job turnover (yearly) \newline Canada \newline 73-86 & Furniture (26.5) \newline Machinery (26.3) \newline Lumber (25.7) & Petroleum (14.1) \newline Primary Metals (13.5) \newline Paper (10.7)  \\
\hline
\citet{Baldwin1994} & Job turnover (yearly) \newline U.S. \newline 73-86 & Lumber (27.2) \newline Apparel (25.5) \newline Leather (22.5) & Petroleum (14.6) \newline Chemicals (14.0) \newline Paper (13.3) \\
\hline   
\end{tabular}\caption{Market Structure measures used, numbers in percent}
\end{table} 
\end{center}

@@

This is a simplified version of the equilibrium conditions in Chen et al. using the notation of Melitz and Ottaviano in which trade freeness is measured by $\rho \in (0,1)$. It captures the main essence of the model, in the short run relative prices depend on the number of firms and negatively on trade freeness (increasing $\rho$ will decrease $\bar{p}$), while in the long country size matters and prices depend positively on trade freeness (increasing $rho$ decreases $1-\rho$ and thus increases $\bar{p}$). Now assume further, that price setting decisions and substitution behavior of consumer is such that we can aggregate \textit{relative} price levels in the same way we aggregated individual price levels before. Then:

@@

\begin{center}
\begin{table*}[ht] \caption{Comparative Statics -- Model Predictions}\label{tb:comp_stats}
\vspace{0.5cm}
{\small
\hfill{}
\begin{tabular}{|l|c|c|c|c|c|c|}
\hline
\textbf{Regressor}& \multicolumn{6}{|c|}{\textbf{Dependent Variables}}\\
\cline{2-7}
\hline
& $P_{sr}$ & $P_{lr}$ & $\mu_{sr}$ & $\mu_{lr}$ & $z_{sr}$ & $z_{lr}$ \\ \hline
$\tau$ & + & -- & + & -- & -- & +\\ \hline
$\tau^*$  & -- & + & -- & + & + & --\\ \hline
$\theta$ & -- & + & -- & + & + & --\\ \hline
$\theta^*$ & + & -- & + & -- & -- & +\\ \hline
$D$  & -- & + & -- & + & + & \\ \hline
$D^*$ & + & -- & + & -- & -- &\\ \hline
$L$ & & & & & & + \\ \hline
$L^*$ & & & & & & -- \\
\hline
\end{tabular}}
\hfill{}
\end{table*}
\end{center}

@@

\bibliography{trade}

>>>>

While these new models of international trade are well grounded in empirical evidence coming from micro data, there are surprisingly few tests of the model predictions for aggregate variables which are decisive for the predicted welfare gains from trade. Recently, \citet{Arkolakisetal2012a,Arkolakisetal2012b} call into question the importance of firm-level heterogeneity by showing that in a lass class of trade models, the additional welfare gains are fairly small and actually even smaller if consumers don't have CES utility. The response of \citet{MelitzRedding2013} shows that there is still considerable disagreement over how to theoretically evaluate the additional welfare gains from firm selection, a fact that motivates us to test the model directly in aggregate data on prices, markups and productivity. To do so, we estimate the effects of trade liberalization on the competitive environment in manufacturing markets of the member countries of the North American Free Trade Agreement (NAFTA). We employ an estimation procedure based on the \citet{MelitzOttaviano2008} model introduced by \citet{Chen2009}, which to our knowledge is the only empirical application of a model with firm-level heterogeneity on aggregate data. \citet{Chen2009} derive estimable regression equations from the model's equilibrium conditions that allow us to test the effects of trade openness on relative price levels, markups and labor productivities of two trading partners. It is further possible to differentiate between the effects of trade in the short run, which, in the model, refers to an economy without relocation decisions for firms, and in the long run, when firms are free to choose their home market for production. However, as the underlying model is static, no direct results on the time path of the impact of trade liberalization can be obtained. We try to address this issue by dividing our sample in ways that make it more amenable to a model-based estimation. Contrary to \citet{Chen2009}, we directly observe tariff rates between the three countries in our sample and hence use those as a direct measure of trade openness. Additionally, we test for the effects of third-country trade openness on the relative performance of two countries that are linked through trade, predictions for which can be derived from the multi-country version of the Melitz and Ottaviano model. Our dataset comprises of nine manufacturing sectors in Canada, Mexico and the US, covering the time period from the introduction of NAFTA in 1994 up to 2006, which gives us reason to believe that we are able to capture the long run effects of policy changes even in industries with low firm churning rates\footnote{Given that the US already engaged in a free-trade agreement with Canada in 1988 (CUSFTA), we also estimate the short-run effects of this agreement prior to NAFTA and the results hold}. \\
Our findings support the main model predictions, with tariff barriers stifling domestic competition, leading to higher producer prices and markups as well as lower productivity. In the immediate years after the free-trade agreement when tariff barriers are reduced, relative prices and markups decrease as relative productivity increases, thus giving rise to competitive effects. The results in the long-run, however, are not as clear cut, with some effects reversing as predicted by the model while some effects persist. This is also confirmed by directly looking at the reaction of industries with different entry barriers to changes in trade openness.  

@@

The paper is organized as follows: Section \ref{sec:lit} gives a survey of the previous literature assessing the effects of trade liberalizations in general and of NAFTA specifically. Section \ref{sec:mo} briefly summarizes the Melitz and Ottaviano (2008) model, derives the most important equilibrium conditions and explains the estimation strategy used in \citet{Chen2009}. Section \ref{sec:app} then presents our application of the model by giving an overview of the data used and our estimation procedure. The results of our regressions and possible shortcomings as well as extensions of our approach are discussed in Section \ref{sec:disc}; Section \ref{sec:conc} concludes.

@@

c_D = \left[ \frac{\phi c_M^k}{L} \frac{1-(\tau^*)^{-k}}{1-(\tau\tau^{*})^{-k}} \right]^{\frac{1}{k+2}} \\
c_D^* = \left[ \frac{\phi c_M^k}{L^*} \frac{1-\tau^{-k}}{1-(\tau\tau^{*})^{-k}} \right]^{\frac{1}{k+2}},

@@

\left( \frac{\bar{p}}{\bar{p}^*} \right)^k = \left( \frac{c_D}{c_D^*} \right)^k = \left(\frac{c_M}{c_M^*} \right)^k \frac{\bar{N}^*}{\bar{N}} \frac{N}{N^*} \frac{1 + \frac{\bar{N}}{\bar{N}^*} \frac{1}{(\tau^*)^k} \left( \frac{c_M^*}{c_M} \right)^k}{1 + \frac{\bar{N}^*}{\bar{N}} \frac{1}{\tau^k} \left( \frac{c_M}{c_M^*} \right)^k} \label{eq:mo-relprice-sr}

@@

\left( \frac{\bar{p}}{\bar{p}^*} \right)^{(k+2)} = \left( \frac{c_D}{c_D^*} \right)^{(k+2)} = \left( \frac{c_M}{c_M^*} \right)^k \frac{L^*}{L} \frac{1-\frac{1}{\tau^k}}{1 - \frac{1}{(\tau^*)^k}} \label{eq:mo-relprice-lr}

@@

These two equations capture one of the central predictions of the Melitz and Ottaviano model: asymmetrical trade liberalizations will have opposing effects on competitiveness in the short and the long run. By equation (\ref{eq:mo-relprice-sr}), lowering trade barriers induces a fall in the cost cutoff, and hence decreases in prices and markups and increases in productivity. In the long run, however, the effects are reversed, as an increase in trade costs induces firms to choose the relatively more protected market for production, thereby increasing competition in markets that are shielded from foreign firms. \\

@@

Similarly, an expression for the average markup can be derived. The determination of the average markup is equivalent to the one for average prices so expressions for the short-- and long--run impacts of openness on markups can readily be derived. Somewhat more problematic is the index for productivity, as the model requires knowledge of a firm's unit costs $c$, which are not observable. Chen et al. work around this issue by assuming away differences in capital costs, so that average industry productivity can be approximated by the ratio of nominal wages to labor productivity: $\bar{c} = \frac{w}{z}$. If it is additionally assumed that unit labor costs only depend on nominal wages, the ratio of domestic to foreign labor productivity can be written as:
\begin{equation}\label{eq:chen-lab-prod}
\frac{z}{z^*} = \frac{w}{w^*} \frac{\bar{c}^*}{\bar{c}}
\end{equation}
If the least competitive firm in an industry with a productivity draw at the upper bound of the distribution $c_M$ has labor productivity $z_M$ and labor is perfectly mobile between firms, equation (\ref{eq:chen-lab-prod}) implies $\frac{z}{z^*} = \frac{w}{w^*} \frac{c_M^*}{c_M}$. This relationship can then be used in an analogous fashion as before to construct an expression relating openness to productivity. In the short run, equation (\ref{eq:chen-lab-prod}) can be amended to yield:
\begin{equation}\label{eq:chen-prod-sr}
\left( \frac{z}{z^*} \right)^k = \left( \frac{z_M}{z_M^*} \right)^k \frac{(\bar{N}/N)}{(\bar{N}^*/N^*)} \frac{1+ \frac{\bar{N}*}{\bar{N}} \frac{\theta}{1-\theta}}{1+ \frac{\bar{N}}{\bar{N}^*} \frac{\theta^*}{1-\theta^*}} 
\end{equation}
Higher values of $\theta$ thus lead to higher productivity (conditional on $\bar{N}/N$), as they force lower productivity firms to shut down production. For the long run, equation (\ref{eq:mo-relprice-sr}) combined with the expression for labor productivity gives:
\begin{equation}\label{eq:chen-prod-lr}
\left( \frac{z}{z*} \right)^{k+2} = \left( \frac{w}{w^*} \right)^2 \frac{L}{L^*} \left( \frac{z_M}{z_M^*} \right)^k \frac{1-\frac{\theta}{1-\theta}}{1-\frac{\theta^*}{1-\theta^*}}
\end{equation} 
Larger markets exhibit higher labor productivity, while the effects of $\theta$ and $\theta^*$ are the opposite of those in the short--run.

@@

\subsection{Further Changes (work in progress!)}
THINK CLOSELY ABOUT THIS PARAGRAPH! \\
Following the arguments in \citet{Trefler2004}, we want to exploit the nature of NAFTA being close to a natural experiment and hence try to identify the effect of the policy measures (i.e. the changes in tariff rates) separately from the effects of trade openness in general. There is evidence that trade openness -- measured by the import penetration of a certain country or industry, as in \citet{Chen2009} -- is affected by a number of external forces, including oil prices \citep{Kilian2009}, credit conditions \citep{Chor2012}, shared culture and language between countries and many more. Therefore, we deconstruct the iceberg costs of trade into two parts: $\tau^{lh}=\frac{Tr^{lh}}{\theta^{lh}}$, where $Tr^{lh}>1$ is the tariff rate for trade between countries $l$ and $h$ and $\theta^{lh}>1$ captures the additional costs of trade imposed by the aforementioned factors. Then, the import penetration can be viewed as a measure of $\theta^{lh}$, as in Chen et al. (2009), while a carefully constructed tariff measure can directly account for the effect of NAFTA-mandated changes in the trade environment. Obviously, the import penetration in a sector depends on the tariff measure as well, so an instrumental variable approach has to be taken in order to identify this effect. We defer the discussion of the construction of the tariff measure and the choice of instruments to the following chapter. Prior work of \citet{Bernard2006} suggests that tariff rates throughout the 1980s, at an average level between four and five percent, accounted for about the same fraction of trade costs as costs directly attached to shipping the good, i.e. freight and insurance, so we expect them to have a sizable impact on trade flows between countries and hence the competitive environment. \vspace{0.5cm} \\

As we have seen in the exposition of the Melitz and Ottaviano model above, there is one crucial caveat in taking the model to the data: due to the static nature of the model, the comparative static results just compare one steady state with another, while being silent about the transitional dynamics. The estimation strategy of \citet{Chen2009} tries to account for this by estimating an error correction model to identify the long-run separately from the short-run, but their results -- just as ours -- are mixed for the long run and it cannot be ruled out that this is due to the estimation procedure. Therefore, we try to address this issue in a more direct way: as short- and long-run in the model differ only in the possibility of firm entry, we separate industries into those with a fixed number of firms and those with low entry barriers. This distinction then gives us industries that represent the short- and long-run and we can directly investigate whether the coefficients on the relevant variables differ significantly\footnote{We were inspired to do so by \citet{Head1999} who use the classification to test competing theories of trade that rely on different market structures.}. This approach, however, leads to two issues that need to be addressed before implementation. First, it is not \textit{a priori} obvious how to measure the entry conditions in an industry; while the theoretical model uses the number of firms, this could in practice either refer to firms or to establishments (i.e. different production sites run by the same parent company), or even to employees, as firms in the model use unit labor input. Second, there is no reason to believe that different measures of entry and exit dynamics are exogenous with respect to trade openness -- indeed in the model trade openness is a key factor in the entry decision of firms, but in the real world there might be various other factors that might lead to industries being asymmetrically affected by a change in trade costs, hence biasing our results. To tackle both these issues, we aim to construct a robust measure of industry dynamics by aggregating multiple studies that examine firm and employment turnover in Canada, Mexico and the United States as well as Europe over different time periods. With this, we hope to identify those industries that are either very dynamic or very static over a broad set of different measures, regions and time periods. Table \ref{tb:market_lit} gives an overview of the studies used and a glance at their respective results, showing considerable variation in the dynamics of entry and job creation in different manufacturing sectors.

\begin{center}
\begin{table}\label{tb:market_lit}
\begin{tabular}{|P{2.8cm}|P{4cm}|P{4cm}|P{4cm}|}
\hline
\textbf{Study}  &  \textbf{Subject}  &  \textbf{Highest Turnover}  & \textbf{Lowest Turnover} \\
\hline 
\hline
\citet{Dunne1988} & Entry Rates (4-yearly) \newline U.S. \newline 63-82    & Instruments (60.3) \newline Lumber (49.70) \newline Printing (49.0)  & Leather (29.4) \newline  Food Processing (23.9) \newline Tobacco (20.5) \\  
\hline
Samianego (2008)   & Entry Rates (yearly)  \newline Europe \newline 97-04  & Paper, printing, software (15.6) \newline Textiles (11.9) \newline Petroleum and Coal (11.9) & Chemicals (9.5) \newline Plastics (9.4) \newline Food Products (9.1) \\
\hline 
\citet{Brown2004}  & Employment renewal    \newline Canada \newline 73-96  & Plastic (79.5) \newline Furniture (79.4) \newline Fabricated Metals (77.2) & Primary Metals (33.6) \newline Paper (32.4) \newline Tobacco (4.2) \\
\hline
\citet{Foster2006} & Job creation (yearly) \newline U.S. \newline 72-98     & Lumber (11.8) \newline Apparel (11.2) \newline Miscellaneous (11.0) & Paper (5.9) \newline Petroleum (5.9) \newline Tobacco (5.1) \\
\hline
\citet{Baldwin1994} & Job turnover (yearly) \newline Canada \newline 73-86 & Furniture (26.5) \newline Machinery (26.3) \newline Lumber (25.7) & Petroleum (14.1) \newline Primary Metals (13.5) \newline Paper (10.7)  \\
\hline
\citet{Baldwin1994} & Job turnover (yearly) \newline U.S. \newline 73-86 & Lumber (27.2) \newline Apparel (25.5) \newline Leather (22.5) & Petroleum (14.6) \newline Chemicals (14.0) \newline Paper (13.3) \\
\hline   
\end{tabular}\caption{Market Structure measures used, numbers in percent}
\end{table} 
\end{center}

@@

This is a simplified version of the equilibrium conditions in Chen et al. using the notation of Melitz and Ottaviano in which trade freeness is measured by $\rho \in (0,1)$. It captures the main essence of the model, in the short run relative prices depend on the number of firms and negatively on trade freeness (increasing $\rho$ will decrease $\bar{p}$), while in the long country size matters and prices depend positively on trade freeness (increasing $\rho$ decreases $1-\rho$ and thus increases $\bar{p}$). Now assume further, that price setting decisions and substitution behavior of consumer is such that we can aggregate \textit{relative} price levels in the same way we aggregated individual price levels before. Then:

@@

\begin{center}
\begin{table*}[ht] \caption{Comparative Statics -- Model Predictions}\label{tb:comp_stats}
\vspace{0.5cm}
{\small
\hfill{}
\begin{tabular}{|l|c|c|c|c|c|c|}
\hline
\textbf{Regressor}& \multicolumn{6}{|c|}{\textbf{Dependent Variables}}\\
\cline{2-7}
\hline
& $\bar{p}_{sr}$ & $\bar{p}_{lr}$ & $\mu_{sr}$ & $\mu_{lr}$ & $z_{sr}$ & $z_{lr}$ \\ \hline
$\tau$     & +  & -- & +  & -- & -- & + \\ \hline
$\tau^*$   & -- & +  & -- & +  & +  & --\\ \hline
$\theta$   & -- & +  & -- & +  & +  & --\\ \hline
$\theta^*$ & +  & -- & +  & -- & -- & + \\ \hline
$D$        & -- &    & -- &    & +  &   \\ \hline
$D^*$      & +  &    & +  &    & -- &   \\ \hline
$L$        &    & -- &    & -- &    & + \\ \hline
$L^*$      &    & +  &    & +  &    & --\\
\hline
\end{tabular}}
\hfill{}
\end{table*}
\end{center}

@@

\bibliography{thesis_bibliography}
